### 2019-09-09 manuscript/manuscript.tex
updates to ms

  Natural selection can proceed in natural populations over short timescales,
  as evident from numerous studies demonstrating rapid phenotypic change.
  However, detecting these changes genome-wide is difficult, as rapid
  adaptation often proceeds from standing variation or is the result of
  polygenic selection, both of which leave faint genomic signals
  indistinguishable from a noisy background of genetic drift. As we developed
  in another paper \parencite{Buffalo2019-io}, one promising signal comes from
  the covariance between allele frequency changes observable from temporal
  genomic data, e.g. evolve-and-resequence studies. These temporal covariances
  reflect how the change in neutral allele frequency at one timepoint is
  predictive of the changes at later timepoints when there is heritable fitness
  variation in the population, as neutral alleles become stochastically linked
  with selected loci. Since genetic drift does not lead to temporal covariance,
  we can begin to directly estimate what fraction of the variation in allele
  frequency change through time is drift by linked selection. Here, we
  re-analyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
  one artificial selection experiment in mice, to quantify the effects of
  linked selection over short timescales. Additionally, taking advantage of
  selection on replicated populations, we can directly estimate the covariance
  in allele frequency between replicates due to convergent selection pressure.
  Overall, we find that in the three studies we analyzed, linked selection has
  a remarkable impact on short term allele frequency dynamics that previously
  may have been attributed to genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles
genetic drift and selection have in shaping genome-wide allele frequency
changes. Selection can both directly and indirectly affect allele frequency
change, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome e.g. linked selection
(\cite{Maynard_Smith1974-lc,Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
focused on teasing apart the impacts of drift and selection on genome-wide
diversity using population samples from a single contemporary timepoint, often
by modeling the correlation between regional recombination rate, gene density,
and diversity created in the presence of linked selection
\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown a major role
for linked selection in shaping patterns of genome-wide diversity across the
genomes of a variety of organisms
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantity the relative roles positive selection
(hitchhiking) and negative selection (background selection) have shaping
patterns of genome-wide diversity
\parencite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
an important dimension: time.

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts that are difficult to distinguish from
background levels of genetic drift and sampling variance. However, increasingly
genomic experimental evolution studies with multiple timepoints, and in some
cases multiple replicate populations, are being used to detect selected loci
\parencite{Turner2011-sx,Turner2012-bm} and differentiate modes of selection
\parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}. In addition these
temporal-genomic studies have begun in wild populations
\parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a previous paper, we
proposed that one useful signal for understanding the genome-wide impact of
polygenic linked selection detectable from temporal genomic studies} is the
temporal autocovariance in allele frequency changes \parencite{Buffalo2019-io}.
These covariances are directly estimable from temporal genomic data and are
created when the loci that underly heritable fitness variation perturb linked
neutral alleles; \vb{in contrast}, when genetic drift acts alone in a closed
population, these covariances are zero in expectation. \vb{Mathematically,
temporal covariances occur because it is natural to decompose the total
variance in allele frequency change across a set of time intervals into the
variances and covariances in allele frequency change among time intervals.
Furthermore, biologically these covariances reflect the extent to which
allele frequency changes in one generation predict changes in another due to
a shared fitness background and shared selection pressures.}

Here, we provide the first empirical analyses quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound of the proportion of total variation in
allele frequency change caused by linked selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change ($\Delta_{_{10}} p_t$) in some initial reference generation $s$ (the row
of the matrix), and some later timepoint $t$ (the column of the matrix). We
corrected these matrices for biases created due to sampling noise, and
normalize the entries for heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and additive genetic variance for fitness
decay \parencite{Buffalo2019-io}. The temporal covariances per replicate are
noisier but this general pattern holds; see Supplementary Figures
\ref{suppfig:barghi-cov-panels}.  \textcite{Barghi2019-qy}'s design means that
the covariances we see in adjacent time intervals are on average ten
generations apart, and given the temporal decay in covariance we see, the
covariances on shorter time-scales may well be higher yet (see Supplementary
Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be decomposed into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is a lower bound on the total effect of linked selection
between the initial generation $0$ and some later generation $t$, which can be
varied to see how this quantity grows through time. As with the temporal
covariances, the study design of \textcite{Barghi2019-qy} leads our measure
$G(t)$ to be even more conservative, since the temporal covariances with each
ten-generation block between sequenced timepoints are not directly observable,
and are not included in the numerator of $G(t)$. Still, we find a remarkably
strong signal that greater than $20\%$ of total variation in allele frequency
change over 60 generations is directly the result of linked selection.

The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
another covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These between-replicate
covariances are created in the same way as temporal covariances are: neutral
alleles linked to a particular fitness background experience are expected to
have allele frequency changes in the same direction if the selection pressures
are similar. We measure this through a statistic similar to a correlation,
which we call the convergent correlation; this is the ratio of average
between-replicate covariance across all pairs to the average standard deviation
across all pairs of replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
quickly, from an initial value of about 0.1 (95\% block bootstrap confidence
intervals $[0.094, 0.11]$), to around 0.01 (95\% CIs [0.0087, 0.015]) within 20
generations.

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other time-points
but the signal of a change in the direction of selection at particular loci may
be washed out when we calculate our genome-wide average temporal covariances.
To address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates).  The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
this}, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the   allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection \vb{and} creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserve the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach).  This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend {\bf XXX ????} (Supplementary Figure
\ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that the
windowed covariances not only decay towards zero, but in fact become negative
through time, consistent with many regions in the genome having had a reversed
fitness effect at later timepoints.

@@

A benefit of between-replicate covariances is unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed allowed us to assess the impact of linked selection
in driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistical significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$). \gc{GIVE conversion \% var
due to linked selection. }

Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
with the two Longshanks comparisons to the control line not being significantly
different from zero, while the comparison between the two Longshanks line is
statistically significantly different from zero (CIs XXX).

@@

We re-analyzed three published temporal genomic datasets and detected the
genome-wide signature of linked selection acting over tens of generations.
Furthermore, we find the dynamics of temporal autocovariance are consistent
with theory; the temporal covariance in allele frequencies decays \vb{as both
  the fitness variation associated with particular haplotypes and the linkage
disequilibria between selected and neutral loci decay through time}. In
reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
generations, greater than 20\% of the variation in allele frequency change is
directly due to the action linked selection. Capitalizing on replicated
evolve-and-resequence study designs, we characterized the extent to which
convergent selection pressures lead to parallel changes in allele frequency
across replicate populations, finding correlation in allele frequency
changes varies from 0.1 (in the \cite{Barghi2019-qy} study) to 0.36 (in the
\cite{Kelly2019-dc} study). \graham{Are these just repeat numbers from main
text? If so nix this sentence}

@@

One limitation of our current work here is that none of the studies we
re-analyzed had linkage disequilibria data for the evolved populations. Our
theory of temporal autocovariance tells us that the temporal autocovariances a
neutral site experiences is determined by the product of the expected linkage
disequilibrium and additive fitness variation that persists through the
generations \parencite{Buffalo2019-ion}. This leads to a clear prediction:
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. Unfortunately, we lack a high-resolution recombination map for
\emph{D. simulans} and while there are LD data for \emph{D. simulans}
\parencite{Signor2018-wg} we did not find a relationship between temporal
covariance and LD. \graham{Is important point. We'll likely need plots for
supp}

@@

Finally, in re-analyzing the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics. This is due either to a change in the
environment and its consequences for the fitness of regions in the genome,
recombination breaking up selected alleles, or epistatic interactions between
loci. It is important to note that the original study did not intentionally
alter the environment in any way; this signal is attained without intentionally
seeking it out. Discerning which of these is creating negative temporal
autocovariance is a sizable challenge, requiring sequencing each generation and
LD data.

@@

It is unclear how strong the fluctuations would have to be to generate a
genome-wide average signal of fluctuating selection from temporal covariances.
For example, many loci could still show a signal of fluctuating selection, but
the average signal could be overwhelmed by other signals of other selection. To
investigate whether there was an genome-wide excess of loci showing evidence of
fluctuating selection we re-analyzed the data of \parencite{Bergland2014-ij}
using the same seasonal fluctuating model as the original paper. This model is
a Binomial logit-linked GLM fit per-locus, where the Spring/Fall seasons are
encoded as a dummy variable, and are regressed on the frequency data. We use
the same binomial weighting procedure as \textcite{Bergland2014-ij}, where the
weights are determined by the effective number of chromosomes, $N_{eff} = (2
n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid
individuals and the read depth at timepoint $t$, respectively). We fit this
model on all loci marked as used in the VCF provided with the
\textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p). Overall, our
p-values for the Wald test for each locus closely match those of the original
paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times 10^{-16}$;
see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison} A), and the
histograms of the p-values are nearly identical (Supplementary Figure
\ref{suppfig:bergland-pvalue-comparison} B). \textcite{Bergland2014-ij} find
loci with a significant association with season after a Benjamini and Hochberg
FDR p-value adjustment \parencite{Benjamini1995-jy}, \vb{however, the null
hypothesis of the Wald test does not give us an idea of the expected number of
variants that may spuriously fit the pattern of seasonal fluctuating
selection.}

>>>>

  Natural selection can proceed in natural populations over short timescales,
  as evident from numerous studies demonstrating rapid phenotypic change.
  However, detecting these changes genome-wide is difficult, as rapid
  adaptation often proceeds from standing variation or is the result of
  polygenic selection, both of which leave faint genomic signals
  indistinguishable from a noisy background of genetic drift. As we developed
  in another paper \parencite{Buffalo2019-io}, one promising signal comes from
  the covariance between allele frequency changes observable from temporal
  genomic data, e.g. evolve-and-resequence studies. These temporal covariances
  reflect how the change in neutral allele frequency at one timepoint is
  predictive of the changes at later timepoints when there is heritable fitness
  variation in the population, as neutral alleles become stochastically linked
  with selected loci. Since genetic drift does not lead to temporal covariance,
  we can begin to directly estimate what fraction of the variation in allele
  frequency change through time is drift by linked selection. Here, we
  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
  one artificial selection experiment in mice, to quantify the effects of
  linked selection over short timescales. Additionally, taking advantage of
  selection on replicated populations, we can directly estimate the covariance
  in allele frequency between replicates due to convergent selection pressure.
  Overall, we find that in the three studies we analyzed, linked selection has
  a remarkable impact on short term allele frequency dynamics that previously
  may have been attributed to genetic drift.

@@

A long-standing problem in evolutionary genetics is quantifying the roles
genetic drift and selection have in shaping genome-wide allele frequency
changes. Selection can both directly and indirectly affect allele frequency
change, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome e.g. linked selection
(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
focused on teasing apart the impacts of drift and selection on genome-wide
diversity using population samples from a single contemporary timepoint, often
by modeling the correlation between regional recombination rate, gene density,
and diversity created in the presence of linked selection
\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach \vb{has shown linked
  selection has a major role in shaping patterns of genome-wide diversity
across the genomes of some of organisms}
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and \vb{negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}}.
However, we lack an understanding of how genome-wide linked selection acts over
an important dimension: time.

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts \vb{on standing variation} that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations,
\vb{with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection} \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
a previous paper, we proposed that one useful signal for understanding the
genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; \vb{in contrast}, when
genetic drift acts alone in a closed population, these covariances are zero in
expectation. \vb{Mathematically, temporal covariances occur because it is
  natural to decompose the total variance in allele frequency change across a
  set of time intervals into the variances and covariances in allele frequency
  change among time intervals.  Furthermore, biologically, these covariances
  reflect the extent to which allele frequency changes in one generation
predict changes in another due to a shared fitness background and shared
selection pressures.}

Here, we provide the first empirical analyses to quantify the impact of linked
selection acting over short timescales (tens of generations) across two evolve
and re-sequence studies \parencite{Barghi2019-qy,Kelly2019-dc}, and one
artificial selection experiment \parencite{Castro2019-uk}. We repeatedly find a
signal of temporal covariance, consistent with linked selection acting to
significantly perturb genome-wide allele frequency changes across the genome in
a manner that other approaches would not be able differentiate from genetic
drift. We estimate the lower bound of the proportion of total variation in
allele frequency change caused by linked selection, and the correlation between
allele frequency changes between replicate populations caused by response to
convergent selection pressures. Overall, we demonstrate that linked selection
has a powerful role in shaping genome-wide allele frequency changes over very
short timescales.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (\vb{in ten generation intervals, denoted} $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
\vb{associated with neutral alleles} decays \parencite{Buffalo2019-io}. The
temporal covariances per replicate are noisier but this general pattern holds;
see Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (\vb{e.g. if
adjacent generations had been sequenced}) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be \vb{partitioned} into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is \vb{this measure of} the effect of linked selection
between the initial generation $0$ and some later generation $t$, which can be
varied to see how this quantity grows through time. As with the temporal
covariances, the study design of \textcite{Barghi2019-qy} leads our measure
$G(t)$ to be even more conservative, since the temporal covariances with each
ten-generation block between sequenced timepoints are not directly observable,
and are not included in the numerator of $G(t)$. Still, we find a remarkably
strong signal that greater than $20\%$ of total variation in allele frequency
change over 60 generations is directly the result of linked selection.

The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
another covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances are: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. \vb{Intuitively, where temporal
  covariances reflect that neutral alleles associated with heritable fitness
  backgrounds are predictive of frequency changes between generations,
  replicate covariances reflect that heritable fitness backgrounds common to
each replicate predict (under the same selection pressures) frequency changes
between replicates.} We measure this through a statistic similar to a
correlation, which we call the convergent correlation: the ratio of average
between-replicate covariance across all pairs to the average standard deviation
across all pairs of replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
quickly, from an initial value of about 0.1 (95\% block bootstrap confidence
intervals $[0.094, 0.11]$), to around 0.01 (95\% CIs [0.0087, 0.015]) within 20
generations.

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
this}, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the   allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection \vb{and} creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach).  This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend \vince{GC: you had ??? here, not sure I see the issue.} (Supplementary
Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a striking pattern that
the windowed covariances not only decay towards zero, but in fact become
negative through time, consistent with many regions in the genome having had a
reversed fitness effect at later timepoints.

@@

A benefit of between-replicate covariances is unlike temporal covariances,
these can be calculated with only two sequenced timepoints and a replicated
study design. This allowed us to assess the impact of linked selection in
driving convergent patterns of allele frequency change across replicate
populations in two other studies. First, we reanalyzed the selection experiment
of \textcite{Kelly2019-dc}, which evolved three replicate wild populations of
\emph{Drosophila simulans} for 14 generations adapting to a novel laboratory
environment. Since each replicate was exposed to the same selection pressure
and share linkage disequilibria common to the original natural founding
population, we expected each of the three replicate populations to have
positive between-replicate covariances. We find all three pairwise
between-replicate covariances are positive and statistical significant (95\%
CIs, [0.0063, 0.0086], [0.0064, 0.0086], [0.0061, 0.0083]). We estimate the
convergent correlation coefficient across these replicates as 0.36 ($95\%$
block-bootstrap confidence interval $[0.31, 0.40]$).

%\gc{TODO GIVE conversion \% var due to linked selection. }

Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (Figure \ref{fig:figure-2} C),
with the two Longshanks comparisons to the control line not being significantly
different from zero, while the comparison between the two Longshanks line is
statistically significantly different from zero (CIs [0.013, 0.037]).

@@

We reanalyzed three published temporal genomic datasets and detected the
genome-wide signature of linked selection acting over tens of generations.
Furthermore, we find the dynamics of temporal autocovariance are consistent
with theory; the temporal covariance in allele frequencies decays \vb{as both
  the fitness variation associated with particular haplotypes and the linkage
disequilibria between selected and neutral loci decay through time}. In
reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
generations, greater than 20\% of the variation in allele frequency change is
directly due to the action linked selection. Capitalizing on replicated
evolve-and-resequence study designs, we characterized the extent to which
convergent selection pressures lead to parallel changes in allele frequency
across replicate populations, finding correlation in allele frequency
changes varies from 0.1 (in the \cite{Barghi2019-qy} study) to 0.36 (in the
\cite{Kelly2019-dc} study). \graham{Are these just repeat numbers from main
text? If so nix this sentence}

@@

One limitation of our current work here is that none of the studies we
reanalyzed had linkage disequilibria data for the evolved populations. Our
theory of temporal autocovariance tells us that the temporal autocovariances a
neutral site experiences is determined by the product of the expected linkage
disequilibrium and additive fitness variation that persists through the
generations \parencite{Buffalo2019-ion}. This leads to a clear prediction:
regions of higher linkage disequilibrium and lower recombination should have
greater temporal autocovariance than regions with lower LD and higher
recombination. Unfortunately, we lack a high-resolution recombination map for
\emph{D. simulans} and while there are LD data for \emph{D. simulans}
\parencite{Signor2018-wg} we did not find a relationship between temporal
covariance and LD. \graham{Is important point. We'll likely need plots for
supp}

@@

Finally, in reanalyzing the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics. This is due either to a change in the
environment and its consequences for the fitness of regions in the genome,
recombination breaking up selected alleles, or epistatic interactions between
loci. It is important to note that the original study did not intentionally
alter the environment in any way; this signal is attained without intentionally
seeking it out. Discerning which of these is creating negative temporal
autocovariance is a sizable challenge, requiring sequencing each generation and
LD data.

@@

It is unclear how strong the fluctuations would have to be to generate a
genome-wide average signal of fluctuating selection from temporal covariances.
For example, many loci could still show a signal of fluctuating selection, but
the average signal could be overwhelmed by other signals of other selection. To
investigate whether there was an genome-wide excess of loci showing evidence of
fluctuating selection we reanalyzed the data of \parencite{Bergland2014-ij}
using the same seasonal fluctuating model as the original paper. This model is
a Binomial logit-linked GLM fit per-locus, where the Spring/Fall seasons are
encoded as a dummy variable, and are regressed on the frequency data. We use
the same binomial weighting procedure as \textcite{Bergland2014-ij}, where the
weights are determined by the effective number of chromosomes, $N_{eff} = (2
n_t d_t - 1) / (2 n_t + d_t)$ ($n_t$ and $d_t$ are the number of diploid
individuals and the read depth at timepoint $t$, respectively). We fit this
model on all loci marked as used in the VCF provided with the
\textcite{Bergland2014-ij} study (doi:10.5061/dryad.v883p). Overall, our
p-values for the Wald test for each locus closely match those of the original
paper (Pearson correlation coefficient 0.98, p-value < $2.2 \times 10^{-16}$;
see Supplementary Figure \ref{suppfig:bergland-pvalue-comparison} A), and the
histograms of the p-values are nearly identical (Supplementary Figure
\ref{suppfig:bergland-pvalue-comparison} B). \textcite{Bergland2014-ij} find
loci with a significant association with season after a Benjamini and Hochberg
FDR p-value adjustment \parencite{Benjamini1995-jy}, \vb{however, the null
hypothesis of the Wald test does not give us an idea of the expected number of
variants that may spuriously fit the pattern of seasonal fluctuating
selection.}
